---
output: 
  rmarkdown::github_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "##",
  fig.path = "images/",
  dpi = 150,
  fig.height = 5,
  fig.width = 10
)
```

# Latent Semantic Scaling

In machine learning, the cost for users to train a model and the control they have over its outputs usually correlate: supervised models are high in control but high in cost; unsupervised models are low in cost but low in control. However, social social scientists have to perform analysis based on existing theories and concepts with very limited computational, financial and human resources.

LSS is a semi-supervised document scaling model developed by Kohei Watanabe to perform large scale analysis of textual data by keeping the balance between the cost and control for [his PhD thesis](http://etheses.lse.ac.uk/3658/). Taking pre-defined *seed words* as weak supervision, it estimates word parameters in a latent semantic space. 

This model has been used for large scale analysis of media content in recent research projects:

* Kohei Watanabe, 2017. "[Measuring News Bias: Russia’s Official News Agency ITAR-TASS’s Coverage of the Ukraine Crisis](http://journals.sagepub.com/eprint/TBc9miIc89njZvY3gyAt/full)", *European Journal Communication*
* Kohei Watanabe, 2017. "[The spread of the Kremlin’s narratives by a western news agency during the Ukraine crisis](http://www.tandfonline.com/eprint/h2IHsz2YKce6uJeeCmcd/full)", *Journal of International Communication*
* Tomila Lankina and Kohei Watanabe. 2017. ["Russian Spring’ or ‘Spring Betrayal’? The Media as a Mirror of Putin’s Evolving Strategy in Ukraine](http://www.tandfonline.com/eprint/tWik7KDfsZv8C2KeNkI5/full)", *Europe-Asia Studies*


## How to install

```{r, eval=FALSE}
devtools::install_github("koheiw/LSS")
```

## How to use

LSS estimates semantic similarity of words based on their surrounding contexts, so a LSS model should be trained on data where the text unit is sentence. It is also affected by noises in data such as function words and punctuation marks, so they should also be removed. It requires larger corpus of texts (5000 or more documents) to accurately estimate semantic proximity. The [sample corpus](https://www.dropbox.com/s/kfhdoifes7z7t6j/data_corpus_guardian2016-10k.RDS?dl=1) contains 10,000 Guardian news articles from 2016.

### Fit a LSS model

```{r, message=FALSE}
require(quanteda)
require(LSS)
```

```{r, cache=TRUE}
corp <- readRDS("/home/kohei/Dropbox/Public/data_corpus_guardian2016-10k.rds")

toks_sent <- corp %>% 
    corpus_reshape("sentences") %>% 
    tokens(remove_punct = TRUE) %>% 
    tokens_remove(stopwords(), padding = TRUE)
dfmt_sent <- toks_sent %>% 
    dfm(remove = "") %>% 
    dfm_select("^\\p{L}+$", valuetype = "regex", min_nchar = 2) %>% 
    dfm_trim(min_termfreq = 5)

eco <- char_keyness(toks_sent, "econom*", p = 0.05)
tmod_lss <- textmodel_lss(dfmt_sent, seedwords('pos-neg'), feature = eco, cache = TRUE)
```

### Sentiment seed words

Seed words are 14 generic sentiment words.

```{r}
seedwords("pos-neg")
```

### Sentiment words

LSS identifies domain specific words and weight them by sentiment.

```{r}
head(coef(tmod_lss), 20) # most positive words
tail(coef(tmod_lss), 20) # most negative words
```

```{r}
par(mar = c(4.1, 10.1, 4.1, 10.1))
plot(tmod_lss$beta, tmod_lss$frequency, type = "n", main = "Word sentiment",
     xlab = "Estimated sentiment", ylab = "Frequency in corpus", log = "y")
text(tmod_lss$beta, tmod_lss$frequency, names(tmod_lss$beta), col = rgb(0, 0, 0, 0.5))
```

## Result of analysis

In the plots, circles indicate sentiment of individual news articles and lines are their local average (solid line) with a confidence band (dotted lines). According to the plot, economic sentiment in the Guardian news stories became negative from February to April, but it become more positive in April. As the referendum approaches, the newspaper's sentiment became less stable, although it became close to neutral (overall mean) on the day of voting (broken line).

```{r}
dfmt <- dfm(corp)

# predict sentiment scores
pred <- as.data.frame(predict(tmod_lss, se.fit = TRUE, newdata = dfmt))
pred$date <- docvars(dfmt, "date")
pred <- na.omit(pred)

# smooth LSS scores
pred_sm <- smooth_lss(pred, from = as.Date("2016-01-01"), to = as.Date("2016-12-31"))

# plot trend
plot(pred$date, pred$fit, col = rgb(0, 0, 0, 0.05), pch = 16, ylim = c(-0.5, 0.5),
     xlab = "Time", ylab = "Negative vs. positive", main = "Economic sentiment in the Guardian")
lines(pred_sm$date, pred_sm$fit, type = "l")
lines(pred_sm$date, pred_sm$fit + pred_sm$se.fit * 2, type = "l", lty = 3)
lines(pred_sm$date, pred_sm$fit - pred_sm$se.fit * 2, type = "l", lty = 3)
abline(h = 0, v = as.Date("2016-06-23"), lty = c(1, 2))
text(as.Date("2016-06-23"), 0.4, "Brexit referendum")
```

