---
title: "README"
author: "KoheiWatanabe"
date: "12 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Latent Semantic Scaling
A highly efficient vector-space model for subject-specific sentiment analysis.

## How to install
```{r, eval=FALSE}
devtools::install_github("koheiw/LSS")
```

## How to use

LSS is created to perform sentiment analysis of long texts, but training of its models should be done on smaller units, typically sentences.

### Fit the model

[Sample dataset](https://www.dropbox.com/s/555sr2ml6wc701p/guardian-sample.RData?dl=0) is a 6000 news articles from the Guardian.

```{r, message=FALSE}
require(quanteda)
require(LSS)
```

```{r}
load('/home/kohei/Dropbox/Public/guardian-sample.RData')

corp_train <- corpus_reshape(data_corpus_guardian, 'sentences')
toks_train <- tokens(corp_train, remove_punct = TRUE)
mt_train <- dfm(toks_train, remove = stopwords())
mt_train <- dfm_remove(mt_train, c('*.uk', '*.com', '*.net', '*.it', '*@*'))
mt_train <- dfm_trim(mt_train, min_count = 10)

#' sentiment model on economy
eco <- char_keyness(toks_train, 'econom*')
lss_eco <- textmodel_lss(mt_train, seedwords('pos-neg'), pattern = eco)
head(lss_eco$beta)
tail(lss_eco$beta)

# sentiment model on politics
pol <- char_keyness(toks_train, 'politi*')
lss_pol <- textmodel_lss(mt_train, seedwords('pos-neg'), pattern = pol)
head(lss_pol$beta)
tail(lss_pol$beta)

```

## Prediction
```{r}
mt <- dfm(data_corpus_guardian)
```

### Economic sentiment

```{r}
sent_eco <- scale(predict(lss_eco, newdata = mt))
plot(docvars(data_corpus_guardian, 'date'), sent_eco, pch = 16, col = rgb(0, 0, 0, 0.1),
     ylim = c(-1, 1), ylab = 'economic sentiment')
lines(lowess(docvars(data_corpus_guardian, 'date'), sent_eco, f = 0.05), col = 1)
abline(h = 0)
```

### Political sentiment

```{r}
sent_pol <- scale(predict(lss_pol, newdata = mt))
plot(docvars(data_corpus_guardian, 'date'), sent_pol, pch = 16, col = rgb(0, 0, 0, 0.1),
      ylim = c(-1, 1), ylab = 'political sentiment')
lines(lowess(docvars(data_corpus_guardian, 'date'), sent_pol, f = 0.05), col = 1)
abline(h = 0)
```

### Comparison

The sentiment analysis models were trained on the same corpus with the same seed, but they are sensitive to different subjects. We can see in the chart below that Guardian's framing of economy became positive in early 2015, while political sentiment were falling down gradually.

```{r}
plot(docvars(data_corpus_guardian, 'date'), rep(0, ndoc(data_corpus_guardian)),  
     type = 'n', ylim = c(-0.5, 0.5), ylab = 'economic/political sentiment')
grid()
lines(lowess(docvars(data_corpus_guardian, 'date'), sent_pol, f = 0.1), col = 1)
lines(lowess(docvars(data_corpus_guardian, 'date'), sent_eco, f = 0.1), col = 2)
abline(h = 0)
legend('topright', lty = 1, col = 1:2, legend = c('political', 'economic'))
```


