---
title: "Application of LSS in research"
subtitle: "Compare Russian state media's coverage of egnergy issues in the US and the EU"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Application of LSS in research}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", 
                      fig.width = 8, fig.height = 4, dpi = 150, out.width = 690)
```

In this example, we analyze how much the proposition of [the European Gas Demand Reduction Plan](https://ec.europa.eu/commission/presscorner/detail/en/ip_22_4608) on 20 July affected Sputnik's coverage of energy issues in the United States and the European Union.

```{r setup}
library(LSX)
library(quanteda)
library(ggplot2)
```

### Preperation

We use the same [corpus](https://www.dropbox.com/s/5tlux53aukf3uqs/data_corpus_sputnik2022.rds?dl=1) as the [introduction](library/LSX/doc/introduction.html), so too the preparation.

```{r include=FALSE}
if (!file.exists("data_corpus_sputnik2022.rds")) {
    download.file("https://www.dropbox.com/s/5tlux53aukf3uqs/data_corpus_sputnik2022.rds?dl=1",
                  "data_corpus_sputnik2022.rds", mode = "wb")
}
```

```{r}
corp <- readRDS("data_corpus_sputnik2022.rds") |>
    corpus_reshape()
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE, 
               remove_numbers = TRUE, remove_url = TRUE)
dfmt <- dfm(toks) |> 
    dfm_remove(stopwords("en"))
```

You need to a [dictionary](https://github.com/koheiw/LSX/blob/master/vignettes/) of keywords used in the analysis.

```{r}
dict <- dictionary(file = "dictionary.yml")
print(dict$keywords)
```


### Estimate the polarity of documents

We can measure the sentiment about energy by collecting words that occur frequently around keywords such as "oil", "gas", "energy" and passing them to `terms`. 

```{r message=FALSE}
seed <- as.seedwords(data_dictionary_sentiment)
term <- char_context(toks, pattern = dict$keywords["energy"], p = 0.01)
lss <- textmodel_lss(dfmt, seeds = seed, terms = term, cache = TRUE, 
                     include_data = TRUE, group_data = TRUE)
```

```{r warning=FALSE}
textplot_terms(lss)
```

### Estimate the polarity of documents

We combine the document variables with the dummy variables in a new a data frame.

```{r}
dat <- docvars(lss$data)
dat$lss <- predict(lss)
print(nrow(dat))
```

### Detect the mentions of country/region 

We can detect mentions of countries using the dictionary. If you want to classify text by country more accurately, you can use the [newsmap package](https://cran.r-project.org/web/packages/newsmap/index.html).

```{r}
dfmt_dict <- dfm(tokens_lookup(toks, dict$keywords[c("us", "eu")]))
print(head(dfmt_dict))
```

We can create dummy variables for mentions of country/region by `dfm_group(dfmt_dict) > 0`. We need to group documents because the unit of analysis is article as `textmodel_lss(group_data = TRUE)` above.

```{r}
mat <- as.matrix(dfm_group(dfmt_dict) > 0)
print(head(mat))
dat <- cbind(dat, mat)
```

### Results

We must smooth the polarity scores of documents separately for the country/region. We can see that difference between the US and EU has expanded soon after the proposition of the European Gas Demand Reduction Plan.

```{r}
smo_us <- smooth_lss(subset(dat, us), lss_var = "lss", date_var = "date")
smo_us$country <- "us"
smo_eu <- smooth_lss(subset(dat, eu), lss_var = "lss", date_var = "date")
smo_eu$country <- "eu"
smo <- rbind(smo_us, smo_eu)
```

```{r}
ggplot(smo, aes(x = date, y = fit, color = country)) + 
    geom_line() +
    geom_ribbon(aes(ymin = fit - se.fit * 1.96, ymax = fit + se.fit * 1.96, fill = country), 
                alpha = 0.1, colour = NA) +
    geom_vline(xintercept = as.Date("2022-06-26"), linetype = "dotted") +
    scale_x_date(date_breaks = "months", date_labels = "%b") +
    labs(title = "Sentiment on egergy (gas and oil)", x = "Date", y = "Sentiment", 
         fill = "Country", color = "Country")
```

To test if the changes after the proposition is statistically significant, we create a dummy variable for the period after the proposition, `after`, and perform regression analysis with its interaction with the country/region dummies. This is akin to [the difference-in-differences design](https://en.wikipedia.org/wiki/Difference_in_differences) commonly used in econometrics.

```{r}
dat_war <- subset(dat, date >= as.Date("2022-02-24"))
dat_war$after <- dat_war$date >= as.Date("2022-06-20")
summary(dat_war[c("lss", "us", "eu", "after")])
```

`dat_war` contains only scores since the beginning the war, so the `intercept` is the average sentiment of articles without mentions of the US or the EU before the proposition; `usTRUE` and `euTRUE` are the average sentiment for the articles with mentions of the US and the EU, respectively.

The coefficient of `afterTRUE` indicates that the overall sentiment became more negative after the proposition (*β* = -0.11; *p* < 0.01). The insignificant coefficient of `euTRUE:afterTRUE` shows that the sentiment for the EU also fell, but the large positive coefficient of `usTRUE:afterTRUE` suggests that the sentiment for the US rose (*β* = 0.22; *p* < 0.001) and became more positive than before.

```{r}
reg <- lm(lss ~ us + eu + after + eu + after + us * after + eu * after, dat_war)
summary(reg) 
```

### Conclusions

Above analysis shows that the Sputnik covered the energy issue in the US more positively while in the EU more negatively after the proposition. Our finding is preliminary, but one possible interpretation is that the Russian government attempted to create divisions between the US and European countries by emphasizing the different impact of the Ukraine war and sanctions against Russia on people's lives.

